Abstraction and Invariance for Algebraically Indexed Types

### 2012-07-09 paper/section-instantiations.tex

\exref{ex:type-iso} is a special case of this isomorphism; another instance is the
type of vector addition:
\begin{eqnarray*}
\forall t_1\mathord:\SynTransl{2}.\forall t_2\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_2}\to\tyPrim{vec}{t_1+t_2}\\
\cong\quad
\forall t_1\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_1}\quad\cong\quad
\tyPrim{vec}{0}
\end{eqnarray*}
\noindent
\newcommand{\transup}[1]{\uparrow_{#1}}
\newcommand{\transdn}[1]{\downarrow_{#1}}
The maps associated with this isomorphism are 
$I = \lambda x. x\ [0]\ 0$ and $J = \lambda y. \Lambda t.\lambda v. \transdn A (y)$,
%\begin{eqnarray*}
%I = \lambda x. x\ [0]\ 0 & \text{ and } &
%J = \lambda y. \Lambda t.\ \lambda v. \trans A v t(y)
%\end{eqnarray*}
where terms $\transdn A : A\to A[0/t]$ and $\transup A : A[0/t] \to A$ are defined by induction over the syntax of $A$:
\begin{eqnarray*}
\transdn{1} & = & \lambda x. x\\ 
\transdn{A_1\times A_2} & = & \lambda p. (\transdn{A_1}(\pi_1 p),\transdn{A_2}(\pi_2 p))\\ 
\transdn{A_1 + A_2} & = & \lambda x.
\textrm{case}\ x\ \textrm{of}\ \textrm{inl}\ y.\textrm{inl}(\transdn{A_1}y)); \textrm{inr}\ y.\textrm{inr}(\transdn{A_2}y)\\ 
\transdn{A_1\to A_2} & = & \lambda f. \lambda x. (\transdn{A_2} (f(\transup{A_1}(x))))\\ 
\transdn{\tyPrim{vec}{e}} & = & \lambda x. x - kv\\ 
\transup{\tyPrim{vec}{e}} & = & \lambda x. x + kv\\ & & \text{ if }e \equiv k t + e'\text{ for
  some }k\in\setOfIntegers \text{ and }e'\text{ with }t\notin e'\\
\end{eqnarray*}
To show $I(J(y))$ is contextually equivalent to $y$ is an easy
induction on $A$ using the fact that the vector $0$
is an identity for vector addition. To show $J(I(x))$ is contextually
equivalent to $x$ we must apply the abstraction theorem and use
the fact that $\tmSem{\transup A}$ and $\tmSem{\transdn A}$ 
considered as relations are simply instances of $\rsem A{}{}$.
%\[
%$\rsem A {t \mapsto v} {} = \{ (x, \tmSem{\transup A (x)}) \mid x \in
%\tySem A\}$.
%\]

>>>>

\exref{ex:type-iso} is a special case of this isomorphism; another instance is the
type of vector addition:
\begin{eqnarray*}
\forall t_1\mathord:\SynTransl{2}.\forall t_2\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_2}\to\tyPrim{vec}{t_1+t_2}\\
\cong\quad
\forall t_1\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_1}\quad\cong\quad
\tyPrim{vec}{0}
\end{eqnarray*}
\noindent
\newcommand{\transup}[1]{\uparrow_{#1}}
\newcommand{\transdn}[1]{\downarrow_{#1}}
The maps associated with this isomorphism are 
$I = \lambda x. x\ [0]\ 0$ and $J = \lambda y. \Lambda t.\lambda v. \transdn A (y)$,
%\begin{eqnarray*}
%I = \lambda x. x\ [0]\ 0 & \text{ and } &
%J = \lambda y. \Lambda t.\ \lambda v. \trans A v t(y)
%\end{eqnarray*}
where terms $\transdn A : A\to A[0/t]$ and $\transup A : A[0/t] \to A$ are defined by induction over the syntax of $A$ (cases omitted for $\transup A$ have definition symmetric to $\transdn A$):
\begin{eqnarray*}
\transdn{1} & = & \lambda x. x\\ 
\transdn{A_1\times A_2} & = & \lambda p. (\transdn{A_1}(\pi_1 p),\transdn{A_2}(\pi_2 p))\\ 
\transdn{A_1 + A_2} & = & \lambda x.
\textrm{case}\ x\ \textrm{of}\ \textrm{inl}\ y.\textrm{inl}(\transdn{A_1}y)); \textrm{inr}\ y.\textrm{inr}(\transdn{A_2}y)\\ 
\transdn{A_1\to A_2} & = & \lambda f. \lambda x. (\transdn{A_2} (f(\transup{A_1}(x))))\\ 
\transdn{\tyPrim{vec}{e}} & = & \lambda x. x - kv\\ 
\transup{\tyPrim{vec}{e}} & = & \lambda x. x + kv\\ & & \text{ if }e \equiv k t + e'\text{ for
  some }k\in\setOfIntegers \text{ and }e'\text{ with }t\notin e'\\
\end{eqnarray*}
To show $I(J(y))$ is contextually equivalent to $y$ is an easy
induction on $A$ using the fact that the vector $0$
is an identity for vector addition. To show $J(I(x))$ is contextually
equivalent to $x$ we must apply the abstraction theorem and use
the fact that $\tmSem{\transup A}$ and $\tmSem{\transdn A}$ 
considered as relations are simply instances of $\rsem A{}{}$.
%\[
%$\rsem A {t \mapsto v} {} = \{ (x, \tmSem{\transup A (x)}) \mid x \in
%\tySem A\}$.
%\]
